Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-22 paper/sections/proof-reconstruction/resolve.tex
[ paper ] table added.
[ paper ] removed HEAD.
[ paper ] fixing lemmas.

Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows equivalence
by converting propositions to their conjunctive normal forms,
and reordering those and the inner disjunctions  to match them
In this section, we provide some theorems to perform such reordering tasks,
the goals are the following.

@@

First, we define the $\rm{rdisj}$ function to convert
a disjunction in its right-associative form.
The $\rm{rconj}$ function that converts conjunction to its right-associative form
is not presented here, but is defined in a similar way (see its implementation
in~\cite{AgdaMetis}).

@@

\begin{lemma}[lem-rdisj]
\label{lem:lem-rdisj}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rdisj}(φ)$.
\end{lemma}

@@

\begin{lemma}[lem-build$_{\boldsymbol{∨}}$]
\label{lem:lem-build-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{build}_{∨}(φ)$.
\end{lemma}

@@

\begin{lemma}[lem-factor]
\label{lem:lem-factor}
  $Γ ⊢ φ$ then $Γ ⊢ \ffactor(φ)$.
\end{lemma}

@@

\begin{lemma}[lem-sbuild$_{\boldsymbol{∨}}$]
\label{lem:lem-sbuild-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{sbuild}_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{theorem}[thm-reorder$_{\boldsymbol{∨}}$]
  \label{thm-reorder-or}
  If $Γ ⊢ φ$ and $ψ: \Prop$ then $Γ ⊢ \freorder_{∨}(φ, ψ)$.
\end{theorem}

@@

\begin{lemma}[lem-reorder$_{\boldsymbol{∧}}$]
  \label{lem:lem-reorder-and}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧}(φ, ψ)$.
\end{lemma}

@@

\begin{lemma}[lem-conjunct$_{\boldsymbol{∨}}$]
  \label{lem:lem-conjunct-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fconjunct_{∨}(φ, ψ)$.
\end{lemma}

@@

\begin{theorem}[thm-reorder$_{\boldsymbol{∧∨}}$]
  \label{thm:thm-reorder-and-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧∨}(φ, ψ)$.
\end{theorem}

@@

\begin{lemma}[lem-rsol]
  \label{lem:lem-rsol}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rsol}(φ)$.
\end{lemma}

@@

\begin{theorem}[thm-resolve]
  \label{thm:thm-resolve}
  If $Γ ⊢ φ₁$, $Γ ⊢ φ₂$, a literal $l$, and $ψ : \Prop$ then
  $Γ ⊢ \fresolve(φ₁, φ₂, l, ψ)$.
\end{theorem}
\Needspace{3\baselineskip}
\begin{proof}
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₁$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
    \vskip 1.5mm
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₂$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
  \vskip 1.5mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{lem-rsol.}
    \UnaryInfC{$Γ ⊢ \rm{rsol}(\freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ))$}
  \end{bprooftree}
  \end{equation*}
\end{proof}

>>>>

Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of the propositions, our approach shows equivalence
by converting propositions to their conjunctive normal forms,
and reordering those and the inner disjunctions to match them.
In this section, we provide some theorems to perform such reordering tasks,
the goals are the following.

@@

First, we define the $\rm{rdisj}$ function to convert
a disjunction in its right-associative form.
The $\rm{rconj}$ function that converts conjunction to its right-associative form
is not presented here, but is defined in a similar way (see its definition
in~\cite{AgdaMetis}).

@@

\begin{mainlemma}[lem-rdisj]
\label{lem:lem-rdisj}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rdisj}(φ)$.
\end{mainlemma}

@@

\begin{mainlemma}[lem-build$_{\boldsymbol{∨}}$]
\label{lem:lem-build-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{build}_{∨}(φ)$.
\end{mainlemma}

@@

\begin{mainlemma}[lem-factor]
\label{lem:lem-factor}
  $Γ ⊢ φ$ then $Γ ⊢ \ffactor(φ)$.
\end{mainlemma}

@@

\begin{mainlemma}[lem-sbuild$_{\boldsymbol{∨}}$]
\label{lem:lem-sbuild-or}
If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \rm{sbuild}_{∨}(φ, ψ)$.
\end{mainlemma}

@@

\begin{mainth}[thm-reorder$_{\boldsymbol{∨}}$]
  \label{thm-reorder-or}
  If $Γ ⊢ φ$ and $ψ: \Prop$ then $Γ ⊢ \freorder_{∨}(φ, ψ)$.
\end{mainth}

@@

\begin{mainlemma}[lem-reorder$_{\boldsymbol{∧}}$]
  \label{lem:lem-reorder-and}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧}(φ, ψ)$.
\end{mainlemma}

@@

\begin{mainlemma}[lem-conjunct$_{\boldsymbol{∨}}$]
  \label{lem:lem-conjunct-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \fconjunct_{∨}(φ, ψ)$.
\end{mainlemma}

@@

\begin{mainth}[thm-reorder$_{\boldsymbol{∧∨}}$]
  \label{thm:thm-reorder-and-or}
  If $Γ ⊢ φ$ and $ψ : \Prop$ then $Γ ⊢ \freorder_{∧∨}(φ, ψ)$.
\end{mainth}

@@

\begin{mainlemma}[lem-rsol]
  \label{lem:lem-rsol}
  If $Γ ⊢ φ$ then $Γ ⊢ \rm{rsol}(φ)$.
\end{mainlemma}

@@

\begin{mainth}[thm-resolve]
  \label{thm:thm-resolve}
  If $Γ ⊢ φ₁$, $Γ ⊢ φ₂$, a literal $l$, and $ψ : \Prop$ then
  $Γ ⊢ \fresolve(φ₁, φ₂, l, ψ)$.
\end{mainth}

\Needspace{3\baselineskip}
\begin{proof}
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₁$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
    \vskip 1.5mm
 \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ₂$}
      \RightLabel{thm-reorder$_{∨}$,}
      \UnaryInfC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
   \end{bprooftree}
  \end{equation*}
  \vskip 1.5mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ)$}
    \AxiomC{$Γ ⊢ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ \freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ)$}
    \RightLabel{lem-rsol.}
    \UnaryInfC{$Γ ⊢ \rm{rsol}(\freorder_{∨}(φ₁, l ∨ ψ) ∧ \freorder_{∨}(φ₂, ¬ l ∨ ψ))$}
  \end{bprooftree}
  \end{equation*}
\end{proof}
